### 2009-04-06 talk/icooolps2009-dotnet/intro.tex
some statements rephrased

Writing a compiler that targets the CLI or JVM is easier than targeting a real CPU, but
it still requires a lot of work, as
IronPython\footnote{\texttt{http://www.codeplex.com/IronPython}},
Jython\footnote{\texttt{http://www.jython.org/}}
and JRuby\footnote{\texttt{http://jruby.codehaus.org/}} demonstrate.

Moreover, writing a static compiler is often not enough to get high
performance.  IronPython and Jython compile code lazily to save memory, but
since they do not exploit the extra information potentially available at
runtime, it is more a delayed static compilation than a true JIT one.
\anto{XXX: someone review the english of the last paragraph, please :-)}

PyPy's approach \cite{RiBo07_223} is to automatize the generation of 
specializing JIT compilers in order
to minimize the effort required to get a fast implementation of a
dynamic language; automatic generation of JIT compilers is done with
the help of partial evaluation techniques and requires the user only
to provide an interpreter with some manual annotations which hint
the generator how interpretation and JIT compilation has to be interleaved \cite{PyPyJIT09}.

>>>>

Writing a compiler that targets the CLI or JVM is easier than targeting a real CPU, but
it still requires a lot of work, as IronPython\footnote{\texttt{http://www.codeplex.com/IronPython}},
Jython\footnote{\texttt{http://www.jython.org/}} and JRuby\footnote{\texttt{http://jruby.codehaus.org/}} demonstrate.
Finally, if one really seeks for an efficent language implementation, \emph{Just In Time} (JIT) compilation needs
to be considered; only in this way  the compiler can exploit runtime type
information to generate quite efficient code. Note that JIT compilation
has not to be confused with lazy compilation of IronPython and Jython which is exploited to save memory, 
since in these cases no runtime type information is ever used to generate more efficient code.

Unfortunately, writing a JIT compiler is a very complex task.  
To make this task simpler, the solution proposed by PyPy \cite{RiBo07_223}  
is automatic generation of JIT compilers with the help of partial evaluation techniques: 
the user has only to provide an interpreter manually annotated with \emph{hints}
specifying how interpretation and JIT compilation has to be interleaved \cite{PyPyJIT09}.
